### 2016-03-02 p00-ch00-intro/intro.tex
Apply comments to intro based on discussion

In a concurrent setting, coroutines~\cite{conway1963design,taocp:knuth} enable interactions with collaborative preemption.
A coroutine allows multiple entry points for suspending and resuming execution at certain locations.
A coroutine instance holds state and the state might vary among different instances.
A coroutine may \emph{yield} to another coroutine;
there's no \emph{return} statement in a coroutine but transfer of control from one to another without returning.
This relation is symmetric and not similar to that of caller-callee type.
This fundamental property makes compositional concurrency of coroutines straightforward.
Coroutines are not originally established in object-oriented principles.
Object orientation focuses on an object-view with caller-callee interaction  whereas coroutines focus on process-view with transfer of control.

In a concurrent setting with objects, multi-threading has been a dominant approach to bring the two together.
A thread holds a single stack of elements.
Execution in a thread is sequential.
The stack has a starting location (i.e. start of invocation) and a finish location (i.e. the return point).
In a caller-callee setting, an object might call methods from other objects.
With objects running in different threads, the execution of methods are interleaved in different threads. 
Future values are used to hold the eventual return value of a method in the caller site.
The unit of interleaving in coroutines is coarse whereas multi-threading is based on more fine-grained scheduling of method invocations.
In addition, interactions in both coroutines and multi-threading are blocking and synchronous.
In contrast, actor model relies on asynchronous communication as one of its core elements.

@@

Considering coroutines, multi-threading, and actor model, it is interesting and challenging to follow the question of how to bring coroutine execution (co-operative scheduling) to an object-oriented setting with asynchronous communication with binding to object interfaces for the messages.

@@

One core challenge is how to create a mapping of coroutines to multi-threading in Java.
Supporting coroutines in Java can be mostly classified in two major categories.
One category relies a \emph{modified} JVM platform (e.g. Da Vinci JVM) in the implementation of thread stack and state such as \cite{Stadler:2009:LCJ:1596655.1596679}, \cite{Stadler:2010:ECJ:1852761.1852765} and \cite{Liu:2006:II:1111320.1111063}.
Other category involves libraries such as Commons Javaflow\footnote{\url{http://commons.apache.org/sandbox/commons-javaflow/}} or Coroutines\footnote{\url{https://github.com/offbynull/coroutines}} that utilize byte-code modification~\cite{dahm1999byte} at runtime in JVM.
In this research, we did not intend to use any of the two above approaches.
We rely on the mainstream JVM platform released by Java team at Oracle; in addition, we do not use byte-code engineering.

Since a thread owns a single stack, a \emph{naive} approach is to map every invocation (entry point) in a coroutine to a new thread in Java. 
This mapping unsurprisingly leads to a poor performance;
the number of threads is not scalable at runtime based on resource limitations in JVM when the number of object increase (cf. Chapter~\ref{ch:p01:ch01}).
Therefore, it is reasonable to use a pool of threads (JSR-166~\cite{jsr166}) to direct the execution of all method invocations (messages). 
However, to deploy a message onto a thread for execution, the message is required to be abstracted in a generic way.
We utilize Java~8 features (JSR 335~\cite{jsr335} and JSR 292~\cite{jsr292:invokedyn}) to abstract a message as a data structure expressed in a lambda expression (cf. Chapter~\ref{ch:p02:ch01}).

@@

\paragraph{Polyglot Programming}
With the rise of distributed computing challenges, software engineering practice has turned to methods that combine multiple programming languages and models to complete a task. 
In this approach, different languages with different focus and abstractions contribute to the same problem statement in different layers. 
Polyglot programming essentially enables software practice to apply the \emph{right} language in the appropriate layer of the problem statement. 
In this research, we aim to deliver ABS semantics and features in a polyglot approach. 
The programmer develops models with ABS that \emph{partially} take advantage of the target language features (e.g. from Java).
This approach is also referred to as \emph{Foreign Function Interface} in the context of ABS modelling.
Listing~\ref{lst:abs:java} shows ABS code that uses \jtt{java.util.ArrayList} as a data strucuture.
With Polyglot programming in ABS, the programmer does \emph{not} need to define the abstractions necessary for the list construct.

\begin{lstlisting}[float=h,language=Java,caption=Using Java in ABS,label=lst:abs:java]
List<String> params = new ArrayList<>();    // Java
myObj ! doSomething(params);                // ABS
\end{lstlisting}

@@

% \paragraph{Extensibility}
% It is a common practice in software engineering that components are developed with a mixture of \emph{configuration}, \emph{extension} and \emph{composition}.
% Hence, delivering the ABS language features as an API brings about a new requirement: to configure or extend the API. 
% In particular, the ABS programmer should be able to extend the API to customized requirements. \TODO{Expand?}

>>>>

In a concurrent setting, coroutines~\cite{conway1963design,taocp:knuth} enable interactions with collaborative preemption.
A coroutine allows multiple entry points for suspending and resuming execution at certain locations.
An instance of a coroutine spawns a new process with its own state.
A coroutine may \emph{yield} to another coroutine;
there's no \emph{return} statement in a coroutine but transfer of control from one to another without returning.
The yield relation is symmetric and not similar to that of caller-callee type.
A coroutine has explicit ability to transfer control to another coroutine.
This fundamental property makes compositional concurrency of coroutines straightforward.
Coroutines are not originally established in object-oriented principles.
Object orientation is based on caller-callee interaction which is asymmetric.
Caller invokes a method in the callee and waits until the \emph{return} occurs in the method of the callee.
Object orientation focuses on an object-view with caller-callee interaction  whereas coroutines focus on process-view with transfer of control.

In a concurrent setting with objects, multi-threading has been a dominant approach to provide an object-oriented concurrency model.
A thread holds a single stack of synchronous method calls.
Execution in a thread is sequential.
The stack has a starting location (i.e. start of invocation) and a finish location (i.e. the return point).
In a caller-callee setting, an object might call methods from other objects.
With objects running in different threads, the execution of methods are interleaved in uncontrollable threads. 
Future values are used to hold the eventual return value of a method in the caller site.
The unit of interleaving in coroutines is coarse whereas multi-threading is based on implicit scheduling of method invocations.
In addition, interactions in both coroutines and multi-threading are blocking and synchronous.
In contrast, actor model relies on asynchronous communication as one of its core elements.

@@

A question naturally rises: how to bring coroutine execution (co-operative scheduling) to an object-oriented setting with asynchronous communication with binding to object interfaces for the messages?
There are a few modelling languages that focus on the question.

Rebeca~\cite{sirjani2002simulation,sirjani2004formal,sirjani2007rebeca} is a modelling language for reactive and concurrent systems. 
In Rebeca, a number of reactive objects (\emph{rebecs}) interact at runtime.
Rebecs interactions is based on asynchronous message passing that leads to a method invocation.
In Rebeca, each rebec has a unique thread of control and an unbounded queue of messages.
Rebeca uses run-to-completion execution and does not support future values.

@@

One core challenge is how to create a mapping of coroutines to multi-threading in Java.
Supporting coroutines in Java can be mostly classified in two major categories.
One category relies a \emph{modified} JVM platform (e.g. Da Vinci JVM) in the implementation of thread stack and state such as \cite{Stadler:2009:LCJ:1596655.1596679}, \cite{Stadler:2010:ECJ:1852761.1852765} and \cite{Liu:2006:II:1111320.1111063}.
Other category involves libraries such as Commons Javaflow\footnote{\url{http://commons.apache.org/sandbox/commons-javaflow/}} or Coroutines\footnote{\url{https://github.com/offbynull/coroutines}} that utilize byte-code modification~\cite{dahm1999byte} at runtime in JVM.
In this research, we did not intend to use any of the two above approaches.
Custom or modified JVM platform implementations are not mainstream and not officially supported by Java team.
Research and development on a modified JVM requires explicit and periodic maintenance and upgrade effort.
Moreover, since byte-code modification changes the byte-code of a running program or inserts new byte-code into JVM at runtime, it makes reasoning and correctness analysis/verification more complicated~\cite{leroy2001java,leroy2003java}.
We rely on the mainstream JVM platform released by Java team at Oracle; in addition, we do not use byte-code engineering.

Since a thread owns a single stack, a straightforward approach is to map every invocation (entry point) in a coroutine to a new thread in Java~\cite{SchaferP10,schafer2010programming}. 
This mapping unsurprisingly leads to a poor performance;
the number of threads is not scalable at runtime based on resource limitations in JVM when the number of object increase (cf. Chapter~\ref{ch:p01:ch01}).
Therefore, it is reasonable to use a pool of threads (JSR-166~\cite{jsr166}) to direct the execution of all method invocations (messages). 
However, to deploy a message onto a thread for execution, the message is required to be abstracted in a generic way.
We utilize Java~8 features (JSR 335~\cite{jsr335} and JSR 292~\cite{jsr292:invokedyn}) to abstract a message as a data structure expressed in a lambda expression (cf. Chapter~\ref{ch:p02:ch01}).

@@

\paragraph{Polyglot Programming}
With the rise of distributed computing challenges, software engineering practice has turned to methods that combine multiple programming languages and models to complete a task. 
In this approach, different languages with different focus and abstractions contribute to the same problem statement in different layers. 
Polyglot programming essentially enables software practice to apply the \emph{right} language in the appropriate layer of the problem statement. 
In this research, we aim to deliver ABS semantics and features in a polyglot approach. 
The programmer develops models with ABS that \emph{partially} take advantage of the target language features (e.g. from Java).
This approach is also referred to as \emph{Foreign Function Interface} in the context of ABS modelling.

\begin{lstlisting}[float=h,language=Java,caption=Using Java in ABS,label=lst:abs:java]
java.util.List<String> params = new java.util.ArrayList<>();    // Java
myObj ! doSomething(params);                                    // ABS
\end{lstlisting}

Listing~\ref{lst:abs:java} shows a snippet of ABS code that uses \jtt{java.util.ArrayList} as a data strucuture.
Ideally in ABS, the programmer is able to directly use the libraries and API from Java.
This removes the necessity to redundantly repeat definition of common data structures and API at ABS.

@@
